### 2022-07-29 paper-code.Rmd
First commit

---
title: "Code for *Increasing Drought Risks amidst Projected Flood Intensification in the Kabul River Basin (Afghanistan and Pakistan) by Khan et al. (2022)*"
author: "Hung Nguyen"
date: "2022-05-14"
output: 
  rmdformats::readthedown: 
    highlight: tango
  pdf_document:
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction and Preparations

This document details the process of producing the results presented in **Increasing Drought Risks amidst Projected Flood Intensification in the Kabul River Basin (Afghanistan and Pakistan) by Khan et al. (2022).

To reproduce the results, please do the following:

-   This code requires R 4.1.0 and above.
-   Download the code repository from the GitHub repo and extract the downloaded `.zip` file to your working folder.
-   Open `chitral-precip.Rproj` in RStudio (**It's important to open this first so that the file path is loaded properly**).
-   Install and load the following packages if you don't already have them. For package `ldsr`, please use the development version which can be installed from GitHub with

```{r, eval=FALSE}
remotes::install_github('ntthung/ldsr')
```

```{r, message=FALSE, warning=FALSE}
library(dplR)       # Tree ring data processing
library(ldsr)       # Tree ring data processing
library(data.table) # Data handling 
library(missMDA)    # Imputation
library(qmap)       # Bias correction
library(modifiedmk) # Trend analysis
library(ggplot2)    # Plotting
library(cowplot)    # Plotting
library(patchwork)  # Plotting
library(ggprism)    # Plotting
theme_set(theme_prism(base_size = 10, base_fontface = 'plain', base_line_size = 0.2))
```

-   Open `paper-code.Rmd`, which is the source code for this document.
-   Follow the written details below and run the code chunks one by one.

For quick access to the final results please see the `.csv` file in `results/`.

The code utilities to support the main code are stored in the folder `R/`. We need to load them first before running the main code.

```{r}
source('R/init.R')
source('R/input-selection-functions.R')
source('R/correlation_functions.R')
source('R/drought_analysis_functions.R')
```

# Data

## Climate data

```{r climate data, warning=FALSE, message=FALSE}
# Monthly temperature
TmWide <- fread('data/chitral-monthly-T.csv')
Tm <- melt(TmWide, id.var = 'year', variable.name = 'month2', variable.factor = TRUE, value.name = 'Tm'
         )[, month := as.integer(month2)
         ][order(year, month)]
# Monthly precipitation
PmWide <- fread('data/chitral-monthly-P.csv')
Pm <- melt(PmWide, id.var = 'year', variable.name = 'month2', variable.factor = TRUE, value.name = 'Pm'
         )[, month := as.integer(month2)
         ][order(year, month)]
```

**Figure 1c**

```{r, fig.width=7, fig.height=3}
p1 <- ggplot(Pm) +
  geom_boxplot(aes(month2, Pm), fill = 'steelblue', alpha = 0.5) +
  labs(x = NULL, y = 'Precipitation [mm]', tag = 'c)') +
  scale_x_discrete(labels = \(x) substr(x, 1, 1)) +
  scale_y_continuous(guide = guide_prism_minor()) +
  panel_border('black', 0.2)

p2 <- ggplot(Tm) +
  geom_boxplot(aes(month2, Tm), fill = 'darkorange', alpha = 0.5) +
  labs(x = NULL, y = 'Temperature [\u00b0C]') +
  scale_x_discrete(labels = \(x) substr(x, 1, 1)) +
  scale_y_continuous(guide = guide_prism_minor()) +
  panel_border('black', 0.2)
p1 + p2
```

## Tree ring data

```{r}
# Read ARSTAN outputs
crnRaw <- lapplyrbind(
  list.files('data/', '.tabs', full.names = TRUE),
  function(fn) {
    dt <- fread(fn)
    dt[, site := substr(fn, 6, nchar(fn) - 5)]
  })
setkey(crnRaw, site)
sssOut <- fread('data/sss.csv', key = 'site')
crn <- merge(crnRaw, sssOut, by = c('site', 'year'))

firstYear <- crn[sss > 0.6][, .N, by = year][N >= 4, first(year)]
lastYear <- 2018
# Use the residual chronology
crnWide <- crn[year %in% firstYear:lastYear, dcast(.SD, year ~ site, value.var = c('res'))]
```

**Figure S1 - Chronology and SSS plot**

```{r, fig.width=7, fig.height=6}
sf <- 2.2
ggplot(sssOut) +
  geom_hline(aes(yintercept = 1, linetype = 'Mean', color = 'Mean')) +
  # geom_hline(aes(yintercept = 0.85 * sf, 
                 # color = 'SSS = 0.85', linetype = 'SSS = 0.85')) +
  geom_hline(aes(yintercept = 0.6 * sf,
                 color = 'SSS = 0.6', linetype = 'SSS = 0.6')) +
  geom_line(aes(year, sss * sf, 
                col = 'Subsample signal strength', 
                linetype = 'Subsample signal strength'), size = 0.5) +
  geom_line(aes(year, res, 
                col = 'Ring width index', 
                linetype = 'Ring width index'), crn) +
  facet_wrap(~site, ncol = 2) +
  scale_x_continuous(
    name = NULL,
    breaks = seq(1300, 2000, 100),
    minor_breaks = seq(1300, 2000, 50),
    guide = guide_prism_minor(),
    expand = c(0, 0)) +
  scale_y_continuous(

    sec.axis = sec_axis(~ . / sf, name = 'Subsample signal strength',
                        breaks = seq(0, 1, 0.25)),
    expand = c(0, 0)) +
  scale_color_manual(
    name = NULL,
    breaks = c('Ring width index', 'Mean',
               'Subsample signal strength', 'SSS = 0.6'),
    values = c('steelblue', 'black', 'darkorange', 'black')) +
  scale_linetype_manual(
    name = NULL,
    breaks = c('Ring width index', 'Mean',
               'Subsample signal strength', 'SSS = 0.6'),
    values = c(1, 1, 1, 2)) +
  guides(color = guide_legend(override.aes = list(size = 0.50))) +
  theme(
    strip.background = element_rect('gray95', NA),
    legend.position = c(0.75, 0.1),
    legend.key.width = unit(1, 'cm'),
    panel.border = element_rect(NA, 'black', 0.2),
    axis.ticks.y.right = element_line(color = 'darkorange'),
    axis.text.y.right = element_text(color = 'darkorange', size = 9),
    axis.title.y.right = element_text(color = 'darkorange', size = 9),
    axis.line.y.right = element_line(color = 'darkorange'),
    axis.ticks.y.left = element_line(color = 'steelblue'),
    axis.text.y.left = element_text(color = 'steelblue', size = 9),
    axis.title.y.left = element_text(color = 'steelblue', size = 9),
    axis.line.y.left = element_line(color = 'steelblue'))
```

## Climate-growth relationship

First we infill the tree ring data

```{r infilling}
X <- crnWide[, -'year'] |> 
  as.matrix() |> 
  imputePCA(ncp = 6) |> 
  {\(x) x$completeObs}() 
PC <- prcomp(X, scale. = TRUE)$x
Xraw <- as.matrix(crnWide[, -'year'])
impModel <- imputePCA(Xraw, ncp = 6)
Xfilled <- impModel$completeObs
Xfitted <- impModel$fittedX
colnames(Xfitted) <- colnames(Xfilled)

crnFitted <- as.data.table(as.data.frame(Xfitted))
crnFitted[, year := 1517:2018]
crnFittedLong <- melt(
  crnFitted, id.vars = 'year', variable.name = 'site', value.name = 'rwi')
crnMerge <- rbindlist(
  list(Infilled = crnFittedLong, Original = crn[, .(site, year, rwi = res)]),
  use.names = TRUE,
  idcol = 'type')
```

**Figure S2**

```{r, fig.width=7, fig.height=7}
ggplot(crnMerge[year >= 1517]) +
  geom_line(aes(year, rwi, color = type)) +
  facet_wrap(vars(site), ncol = 2, scales = 'free_y') +
  scale_color_manual(values = c('darkorange', 'steelblue')) +
  scale_x_continuous(breaks = seq(1500, 2000, 100)) +
  labs(x = NULL, y = 'RWI [-]') +
  panel_border('black', 0.2) +
  theme(legend.position = c(0.6, 0.1))
```

Calculate bootstrapped correlations

```{r}
treeYears <- crnWide$year
# Current year
# Merge TR in 1965-2018 with precipitation in 1965-2018
instIndc <- which(treeYears %in% 1965:2018) 
XYc <- cbind(Xfilled[instIndc, ], as.matrix(PmWide[, -'year']))

# Tree rings and previous year precipitation
# Merge TR in 1966-2018 with precipitation in 1965-2017
instIndp <- which((treeYears - 1) %in% 1965:2017) # previous year streamflow
XYp <- cbind(Xfilled[instIndp, ], as.matrix(PmWide[-.N, -'year']))

# Tree rings and next year precipitation
# Merge TR in 1963-2016 with precipitation in 1965-2018
# instIndn <- which((treeYears + 1) %in% 1965:2018) # next year streamflow
# XYn <- cbind(Xfilled[instIndn, ], as.matrix(PmWide[, -'year']))

set.seed(2022)
corDTc <- cor_boot(XYc, 1:7, 8:19, groupNames = c('site', 'month')) # current year
corDTp <- cor_boot(XYp, 1:7, 8:19, groupNames = c('site', 'month')) # current year
# corDTn <- cor_boot(XYn, 1:7, 8:19, groupNames = c('site', 'month')) # current year

corDTc[, month := paste0(month, 'c')]
corDTp[, month := paste0(month, 'p')]
# corDTn[, month := paste0(month, 'n')]

corDTPm <- rbind(corDTp, corDTc)
corDTPm[, month := factor(month, c(paste0(month.abb, 'p'), paste0(month.abb, 'c')))]
```

**Figure 2**

```{r, fig.width=7, fig.height=6}
ggplot(corDTPm) +
  geom_linerange(
    aes(x = month, ymin = low, ymax = high, alpha = signif), color = 'steelblue') +
  geom_point(
    aes(x = month, y = rho0, alpha = signif), color = 'steelblue') +
  geom_hline(aes(yintercept = 0)) +
  geom_vline(xintercept = 12.5) +
  facet_wrap(vars(site), ncol = 2) +
  scale_x_discrete(name = NULL, labels = monthLab) +
  scale_y_continuous(name = 'Correlation [-]', 
                     breaks = c(-0.6, -0.4, -0.2, 0, 0.2, 0.4, 0.6)) +
  scale_alpha_manual(
    values = c(0.25, 1), labels = c('Not significant', 'Significant')) +
  theme(
    legend.position = c(0.75, 0.1),
    axis.text.x = ggtext::element_markdown()) +
  panel_border('black', 0.2)
```

# September to August precipitation reconstruction

## Principal compnent analysis

```{r}
PCfit <- prcomp(X, scale. = TRUE)
PC <- PCfit$x
pcDT <- as.data.table(as.data.frame(summary(PCfit)$importance),
                      keep.rownames = 'var') |> 
  melt(id.vars = 'var', variable.name = 'PC')
```

**Figure S3**

```{r, fig.width=7, fig.height=4}
p1 <- ggplot(pcDT[var == 'Standard deviation']) +
  geom_col(aes(PC, value), color = 'black', fill = 'gray95') +
  geom_hline(yintercept = 1, size = 0.8) +
  scale_y_continuous(expand = expansion(add = c(0, 0.01)),
                     limits = c(0, 2)) +
  labs(x = NULL, y = 'Eigenvalue [-]', tag = 'a)') +
  theme(
    panel.grid.major.y = element_line('gray'),
    panel.ontop = TRUE,
    axis.ticks = element_blank(),
    axis.line = element_blank())
p2 <- ggplot() +
  geom_col(
    aes(PC, value * 100), 
    pcDT[var == 'Proportion of Variance'],
    fill = 'steelblue', color = 'black', alpha = 0.8) +
  geom_line(
    aes(PC, value * 100, group = 1),
    pcDT[var == 'Cumulative Proportion'],
    color = 'firebrick', size = 0.8) +
  geom_point(
    aes(PC, value * 100),
    pcDT[var == 'Cumulative Proportion'],
    color = 'firebrick', size = 1.6) +
  labs(x = NULL, y = 'Proportion of variance [%]', tag = 'b)') +
  scale_y_continuous(expand = expansion(add = c(0, 1))) +
  theme(
    panel.grid.major.y = element_line('gray'),
    panel.ontop = TRUE,
    axis.ticks = element_blank(),
    axis.line = element_blank())

p1 + p2
```

## Reconstruction

```{r}
Pm[, year2 := fifelse(month %in% 9:12, year + 1, year)]
PSepAug <- Pm[, .(Qa = sum(Pm)), by = .(year = year2)][year %in% 1966:lastYear]
instInd <- which(firstYear:lastYear %in% PSepAug$year)

sv <- c(1, 2)
sPC <- as.data.table(PC[, sv, drop = FALSE])
lmFit <- PCR_reconstruction(PSepAug, sPC, firstYear, transform = 'none')
set.seed(42)
cvFolds <- make_Z(PSepAug$Qa, frac = 0.25, contiguous = TRUE)
lmCV <- cvPCR(PSepAug, sPC, firstYear, transform = 'none',
              Z = cvFolds, use.robust.mean = FALSE)
```

Performance scores

```{r}
round(lmCV$metrics, 2)
```
**Figure S4 - score distribution**

```{r, fig.width=7, fig.height=5}
scores <- melt(lmCV$metrics.dist[, .(R2, RE, CE)], 
               measure.vars = 1:3,
               variable.name = 'metric')
ggplot(scores) +
  geom_boxplot(aes(metric, value), fill = 'grey95') +
  geom_jitter(aes(metric, value, color = metric), width = 0.25) +
  scale_color_brewer(palette = 'Set2') +
  scale_x_discrete(
    name = NULL,
    guide = guide_prism_bracket(),
    labels = c('R\u00b2', 'RE', 'CE')) +
  scale_y_continuous(
    name = 'Metric value [-]',
    limits = c(0.25, 0.85),
    breaks = seq(0.2, 0.9, 0.1),
    minor_breaks = seq(0.25, 0.85, 0.05),
    guide = guide_prism_offset_minor()) +
  theme(legend.position = 'none')
```

Final reconstruction with bias correction

```{r}
recFinal <- lmFit$rec
recFinal[, lp20 := dplR::pass.filt(Q, 20, 'low')]
recFinal[, lp50 := dplR::pass.filt(Q, 50, 'low')]
bcFit <- fitQmap(
  PSepAug$Qa,
  recFinal[year %in% PSepAug$year, Q],
  'RQUANT',
  wet.day = FALSE)
recFinal[, Qbc := doQmap(Q, bcFit)]
```

**Figure 3**

```{r, fig.width=7, fig.height=4}
p1 <- ggplot(recFinal[year %in% PSepAug$year]) +
  geom_line(aes(year, Q, colour = 'Reconstruction', linetype = 'Reconstruction'), size = 0.4) +
  geom_line(aes(year, Qbc, colour = 'Bias-corrected', linetype = 'Bias-corrected'), size = 0.4) +
  geom_line(aes(year, Qa, colour = 'Observation', linetype = 'Observation'), 
            PSepAug, size = 0.4) +
  scale_colour_manual(
    name = NULL, 
    values = c('Observation' = 'gray30',
               'Reconstruction' = 'steelblue',
               'Bias-corrected' = 'darkorange')) +
  scale_linetype_manual(
    name = NULL, 
    values = c('Observation' = 1, 
               'Reconstruction' = 1,
               'Bias-corrected' = 2)) +
  scale_x_continuous(
    minor_breaks = seq(1965, 2020, 5),
    limits = c(1966, 2018), 
    guide = guide_prism_offset_minor()) +
  scale_y_continuous(
    minor_breaks = seq(100, 850, 50),
    breaks = seq(100, 850, 100),
    limits = c(100, 850),
    guide = guide_prism_offset_minor()) +
  labs(x = 'Year', y = 'Precipitation [mm]', tag = 'a)') +
  theme(
    plot.margin = margin(r = 10),
    legend.box.margin = margin(),
    legend.margin = margin(),
    legend.key.width = unit(2, 'cm'),
    legend.position = 'top') 
p2 <- ggplot() +
  stat_density(aes(y = Q, colour = 'Reconstruction', linetype = 'Reconstruction'), 
               recFinal[year %in% PSepAug$year], geom = 'line', bw = 60, size = 0.4) +
  stat_density(aes(y = Qbc, colour = 'Bias-corrected', linetype = 'Bias-corrected'), 
               recFinal[year %in% PSepAug$year], geom = 'line', bw = 70, size = 0.4) +
  stat_density(aes(y = Qa, colour = 'Observation', linetype = 'Observation'), 
               PSepAug, geom = 'line', bw = 70, size = 0.4) +
  scale_colour_manual(
    name = NULL, 
    values = c('Observation' = 'gray30',
               'Reconstruction' = 'steelblue',
               'Bias-corrected' = 'darkorange')) +
  scale_linetype_manual(
    name = NULL, 
    values = c('Observation' = 1, 
               'Reconstruction' = 1,
               'Bias-corrected' = 2)) +
  scale_x_continuous(guide = guide_prism_offset()) +
  scale_y_continuous(
    minor_breaks = seq(100, 850, 50),
    breaks = seq(100, 850, 100),
    limits = c(100, 850),
    guide = guide_prism_offset_minor()) +
  labs(y = NULL, x = 'Density', colour = NULL, tag = 'b)') +
  theme(
    axis.text.y = element_blank(),
    legend.box.margin = margin(),
    legend.margin = margin(),
    legend.key.width = unit(2, 'cm'),
    legend.position = 'top') +
  theme(legend.position = 'top')

layout <- '
CCC
AAB
AAB
AAB
AAB
AAB
AAB
'

p1 + p2 + guide_area() +
  plot_layout(design = layout, guides = 'collect')
```

# Trend and drought analyses

Calculate rolling statistics and extract drought events

```{r}
recFinal[, rolMax := frollapply(Qbc, 50, max, align = 'right')]
recFinal[, rolMin := frollapply(Qbc, 50, min, align = 'right')]
recFinal[, rolMed := frollapply(Qbc, 50, median, align = 'right')]
recFinal[, period := fcase(
  year %in% 1517:1767, 1,
  default = 2)]
densCals <- recFinal[, {
  d <- density(Qbc, cut = 0, bw = 45)
  list(x = d$x, y = d$y / max(d$y) * 1.5)
}, by = period]

medians <- recFinal[, {
  d <- density(Qbc, cut = 0, bw = 45)
  m <- median(Qbc)
  y <- approx(d$x, d$y, m)$y
  list(x = m, y = y / max(d$y) * 1.5)
}, by = period]

recFinal[, dP := Qbc - mean(Qbc),
       ][, type := classify_events(dP)
       ][, dp10 := pass.filt(dP, 10)]
droughts <- get_timing(recFinal$dP, recFinal$type)[type == 'drought']
droughts[, ':='(yearStart = recFinal[start, year],
                period = recFinal[start, period],
                yearFinal = recFinal[final, year])]
worstDroughts <- droughts[order(peak)][1:3]
```

Mann-Kendall trend test with trend-free prewhitening

```{r}
recFinal[!is.na(rolMax), tfpwmk(rolMax)] |> round(6)
recFinal[!is.na(rolMin), tfpwmk(rolMin)] |> round(6)
recFinal[!is.na(rolMed), tfpwmk(rolMed)] |> round(6)
```

**Figure 4**

```{r, fig.width=8, fig.height=5, message=FALSE}
p1 <- ggplot(recFinal) +
  geom_line(aes(year, Qbc, color = 'Annual')) +
  labs(x = NULL, y = 'P [mm]') +
  geom_line(aes(year, rolMax, colour = '50-yr max'), na.rm = TRUE, size = 0.4) +
  geom_line(aes(year, rolMin, colour = '50-yr min'), na.rm = TRUE, size = 0.4) +
  geom_line(aes(year, rolMed, colour = '50-yr median'), na.rm = TRUE, size = 0.4) +
  geom_smooth(aes(year, rolMax, colour = '50-yr max'), size = 0.6,
              formula = 'y ~ x', method = 'lm', na.rm = TRUE, fill = NA) +
  geom_smooth(aes(year, rolMin, colour = '50-yr min'), size = 0.6,
              formula = 'y ~ x', method = 'lm', na.rm = TRUE, fill = NA) +
  geom_smooth(aes(year, rolMed, colour = '50-yr median'), size= 0.6,
              formula = 'y ~ x', method = 'lm', na.rm = TRUE, fill = NA) +
  scale_x_continuous(
    expand = c(0, 0),
    minor_breaks = seq(1500, 2025, 25),
    limits = c(1500, 2025),
    guide = guide_prism_offset_minor()) +
  scale_y_continuous(
    minor_breaks = seq(200, 900, 50),
    breaks = seq(200, 900, 100),
    limits = c(200, 900),
    labels = skip_label(2),
    guide = guide_prism_offset_minor()) +
  labs(x = NULL, y = 'Precipitation [mm]', colour = NULL, tag = 'a)') +
  scale_color_manual(values = c(RColorBrewer::brewer.pal(3, 'Set2'), 'gray')) +
  theme(
    legend.key.width = unit(0.5, 'cm'),
    legend.position = 'top')

p2 <- ggplot(densCals) +
  geom_ribbon(aes(x, ymin = period, ymax = y + period, group = factor(period),
                  fill = factor(period)), 
              alpha = 0.5) +
  geom_line(aes(x, y + period, group = factor(period))) +
  geom_linerange(aes(x, ymin = period, ymax = period + y), medians, colour = 'firebrick') +
  scale_fill_manual(
    labels = c('1517-1717', '1718-2018'),
    values = wesanderson::wes_palette('Royal1', 4)[c(1, 4)]) +
  scale_x_continuous(
    guide = guide_prism_offset_minor(),
    breaks = seq(200, 900, 100), 
    minor_breaks = seq(200, 900, 20)) +
  scale_y_continuous(
    expand = c(0, 0),
    breaks = c(1.5, 2.5),
    labels = c('1517-1717', '            1718-2018')) +
  labs(x = 'Precipitation [mm]', y = NULL, tag = 'b)') +
  theme(
    legend.position = 'none',
    axis.line.y = element_blank(),
    axis.text.y.left = element_text(angle = 90, hjust = 0.5),
    axis.ticks.y = element_blank())

p3 <- ggplot(recFinal) +
  geom_rect(
    aes(xmin = yearStart, xmax = yearFinal, ymin = -Inf, ymax = Inf),
    worstDroughts,
    fill = 'magenta4', alpha = 0.1) +
  geom_col(aes(year, dP, fill = dP)) +
  scale_fill_distiller(
    palette = 'BrBG', limits = abs_range(recFinal$dP),
    guide = guide_none(),
    direction = 1) +
  scale_color_manual(values = 'black') +
  scale_x_continuous(
    expand = c(0, 0),
    minor_breaks = seq(1500, 2025, 25),
    limits = c(1500, 2025),
    guide = guide_prism_offset_minor()) +
  scale_y_continuous(
    breaks = seq(-300, 400, 100),
    minor_breaks = seq(-300, 400, 50),
    guide = guide_prism_offset_minor()) +
  labs(x = NULL, y = 'Precipitation anomaly [mm]', tag = 'c)') +
  theme(
    legend.position = 'top')

p4 <- ggplot(droughts) +
  geom_point(aes(dur, peak, color = as.character(period)), alpha = 0.6) +
  geom_text(
    aes(dur, peak, label = paste(yearStart, yearFinal, sep = '-')), 
    worstDroughts,
    size = 2, hjust = 0.8, nudge_y = -15) +
  # scale_x_continuous(
  #   breaks = seq(2, 20, 2),
  #   guide = guide_prism_offset_minor()) +
  # scale_y_continuous(
  #   breaks = seq(-300, 0, 50),
  #   limits = c(-275, -25),
  #   guide = guide_prism_offset_minor()) +
  scale_color_manual(
    labels = c('1517-1717', '1718-2018'),
    values = wesanderson::wes_palette('Royal1', 4)[c(1, 4)]) +
  labs(x = 'Drought duration [years]', y = 'Drought severity [mm]', tag = 'd)') +
  theme(legend.position = 'top')

p1 + p2 + p3 + p4 +
  plot_layout(widths = c(1.5, 1))
```

Export the reconstruction

```{r export, eval=FALSE}
recFinal[, period := NULL]
fwrite(
  recFinal[, .(year, P = Q, Plower = Ql, Pupper = Qu, Pbc = Qbc)], 
  'results/chitral-sep-aug-precip-reconst.csv')
```

### 2022-07-29 R/correlation_functions.R

@@ -1,0 +1,134 @@
+#' Get the boundaries of areas of significant correlation
+#' 
+#' Returns columsn of top, down, bottom, left, TRUE for that respective border
+signif_area <- function(cor.dt, dx, dy) {
+  
+  cor.dt[{signif}, list(long, lat,
+                        top    = .SD[.(long,      lat + dy), is.na(signif)],
+                        bottom = .SD[.(long,      lat - dy), is.na(signif)],
+                        left   = .SD[.(long - dx, lat     ), is.na(signif)],
+                        right  = .SD[.(long + dx, lat     ), is.na(signif)])]
+}
+
+#' Get correlatin matrix between two groups of variables.
+#' 
+#' @param x A matrix, with variable names encoded in column names.
+#' @param colInd1 An integer or character vector containing the indices or column names of the first group.
+#' @param colInd2 An integer or character vector containing the indices or column names of the second group.
+#' @return A correlation matrix as returned by `cor()`.
+
+cor_mat <- function(x, colInd1, colInd2) {
+  cor(x[, colInd1], x[, colInd2], use = 'pairwise.complete.obs')
+}
+
+#' Bootstrapped correlation
+#' 
+#' Calculate the observed correlation and boostrap correlation quantiles
+#' @inheritParams cor_mat
+#' @param B Number of bootstrap replicates
+#' @param alpha Significance level, e.g. 0.05, 0.1, etc (must be less than 0.5).
+#' @param groupNames A character vector of length 2: the names of the two groups of variables, in that order.
+#' @return A data.table with seven columns: 
+#' * The first two columns are the two groups, each row is a combination of variables.
+#' * The next three columns are correlation results: low (alpha-th percentile), median, and high ((1-alpha)-th percentile).
+#' * signif, which is TRUE if 0 is not between (low, high).
+#' * rho0: the sampling correlation, which is the `t0` returned by `boot::tsboot()`.
+
+cor_boot <- function(X, colInd1, colInd2, groupNames, B = 1000, alpha = 0.05) {
+  stopifnot(alpha < 0.5)
+  N <- nrow(X)
[…]

### 2022-07-29 R/drought_analysis_functions.R

@@ -1,0 +1,85 @@
+# Classify each year
+classify_events <- function(x) {
+  N <- length(x)
+  events <- rep(0, N)
+  if (x[1] < 0 && x[2] < 0) {
+    events[1] <- events[2] <- -1
+  } else if (x[1] > 0 && x[2] > 0) {
+    events[1] <- events[2] <- 1
+  }
+  for (i in 2:(N-1)) {
+    if (events[i-1] == -1) {             # In drought
+      if (x[i] < 0) {                    # This year is dry: still in drought
+        events[i] <- -1
+      } else {                           # This year is wet
+        if (x[i+1] < 0) {                # Next year back to dry: still in drought
+          events[i] <- -1
+        } else {                         # Next year also wet: turn to pluvial
+          events[i] <- 1
+        }
+      }
+    } else {                             # Not in drought
+      if (events[i-1] == 1) {            # In pluvial
+        if (x[i] > 0) {                  # This year is wet: still in pluvial
+          events[i] <- 1
+        } else {                         # This year is dry
+          if (x[i+1] > 0) {              # Next year back to wet: still in pluvial
+            events[i] <- 1
+          } else {                       # Next year is also dry: turn to drought
+            events[i] <- -1
+          }
+        }
+      } else {                           # In neutral
+        if (x[i] < 0 && x[i+1] < 0) {    # Start a drought
+          events[i] <- -1
+        } else {
+          if (x[i] > 0 && x[i+1] > 0) {  # Start a pluvial
+            events[i] <- 1
+          }                              # Otherwise stay neutral
+        }
+      }
+    }
+  }
+  if (events[N-1] == -1 && x[N] < 0) {
+    events[N] <- -1
+  } else {
+    if (events[N-1] == 1 && x[N] > 0) {
+      events[N] <- 1
+    }
+  }
+  events
+}
+
+# Make a data.table of each events, consisting of timing, duration, and peak
+get_timing <- function(x, events) {
+  
+  N <- length(events)
+  
+  # Get start and end of sequences
+  events <- c(0, events)
+  droughtStt <- which(sapply(2:N, \(k) events[k-1] != -1 && events[k] == -1))
+  pluvialStt <- which(sapply(2:N, \(k) events[k-1] !=  1 && events[k] ==  1))
[…]

### 2022-07-29 R/init.R

@@ -1,0 +1,61 @@
+update_geom_defaults('line', list(size = 0.2))
+update_geom_defaults('hline', list(size = 0.2))
+update_geom_defaults('vline', list(size = 0.2))
+
+# Labeling
+skip_label <- function(n) {
+  function(x) {
+    idx <- seq_along(x)
+    x[idx %% n != 1] <- ' '
+    x 
+  }
+}
+
+monthLab <- function(x) {
+  sapply(x, function (m) {
+    char <- substr(m, 4, 4)
+    mon <- substr(m, 1, 1)
+    fcase(char == 'p', paste0('<span style="color:firebrick">', mon, '</span>'),
+          char == 'n', paste0('<span style="color:darkorange">', mon, '</span>'),
+          default = mon)
+  })
+}
+
+# Converting dplR objects to data.table
+
+rwl_to_dt <- function(rwl, var.name = 'core', val.name = 'rw') {
+  years <- as.integer(row.names(rwl))
+  DT <- as.data.table(rwl)
+  DT[, year := years]
+  DT <- melt(DT, id.vars = 'year', variable.name = var.name, value.name = val.name)
+  DT[!is.na(get(val.name))]
+}
+
+crn_to_dt <- function(crn) {
+  DT <- as.data.table(crn, keep.rownames = 'year')
+  DT[, year := as.integer(year)]
+  DT[]
+}
+
+# Other utilities
+
+lapplyrbind <- function(x, fun, ..., id = NULL) rbindlist(lapply(x, fun, ...), idcol = id)
+
+abs_range <- function(x) {
+  absMax <- max(abs(x))
+  c(-absMax, absMax)
+}
+
+standardize <- function(x, ...) (x - mean(x, ...)) / sd(x, ...)
+
+roundDT <- function(DT, digits = 2, type = 'numeric') {
+  DT[, lapply(.SD, 
+              function(x) {
+                if (is.numeric(x)) {
+                  rd <- round(x, digits)
+                  if (type == 'numeric') rd else sprintf(glue("%.{digits}f"), rd)
+                } else x
+              })]
+}
+
+`%ni%` <- Negate(`%in%`)

### 2022-08-03 paper-code.Rmd
Update title, revise Figure 4 so that it is more legible, and add minor edits.

@@
 ---
-title: "Code for *Increasing Drought Risks amidst Projected Flood Intensification in the Kabul River Basin (Afghanistan and Pakistan) by Khan et al. (2022)*"
+title: "Code for *Tree-Ring Evidence of Increasing Drought Risks amidst Projected Flood Intensification in the Kabul River Basin (Afghanistan and Pakistan)* by Khan et al. (2022)"
 author: "Hung Nguyen"
@@
 output: 
-  rmdformats::readthedown: 
-    highlight: tango
-  pdf_document:
-    highlight: tango
+  rmdformats::readthedown: default 
+  pdf_document: default
 ---
@@
 
-This document details the process of producing the results presented in **Increasing Drought Risks amidst Projected Flood Intensification in the Kabul River Basin (Afghanistan and Pakistan) by Khan et al.
+This document details the process of producing the results presented in **Tree-Ring Evidence of Increasing Drought Risks amidst Projected Flood Intensification in the Kabul River Basin (Afghanistan and Pakistan)** by Khan et al.
 (2022).
@@
 library(ggprism)    # Plotting
+library(ggnewscale) # Plotting
 theme_set(theme_prism(base_size = 10, base_fontface = 'plain', base_line_size = 0.2))
@@
 -   Follow the written details below and run the code chunks one by one.
+
+This R Markdown is set to render both HTML and PDF outputs. To do so, please run
+
+```{r, eval=FALSE}
+rmarkdown::render('paper-code.Rmd', output_format = 'all', output_options = list(hightlight = 'tango'))
+```
 
@@
 ```{r}
-recFinal[, rolMax := frollapply(Qbc, 50, max, align = 'right')]
-recFinal[, rolMin := frollapply(Qbc, 50, min, align = 'right')]
-recFinal[, rolMed := frollapply(Qbc, 50, median, align = 'right')]
+alignType <- 'right'
+recFinal[, rolMax := frollapply(Qbc, 50, max, align = alignType)]
+recFinal[, rolMin := frollapply(Qbc, 50, min, align = alignType)]
+recFinal[, rolMed := frollapply(Qbc, 50, median, align = alignType)]
 recFinal[, period := fcase(
@@
 
-```{r, fig.width=8, fig.height=5, message=FALSE}
+```{r, fig.width=8, fig.height=6, message=FALSE}
 p1 <- ggplot(recFinal) +
+  labs(x = NULL, y = 'P [mm]') +
+  geom_line(aes(year, rolMax, colour = '50-yr max'), na.rm = TRUE, size = 0.4, alpha = 0.5) +
+  geom_line(aes(year, rolMin, colour = '50-yr min'), na.rm = TRUE, size = 0.4, alpha = 0.5) +
+  geom_line(aes(year, rolMed, colour = '50-yr median'), na.rm = TRUE, size = 0.4, alpha = 1) +
   geom_line(aes(year, Qbc, color = 'Annual')) +
-  labs(x = NULL, y = 'P [mm]') +
-  geom_line(aes(year, rolMax, colour = '50-yr max'), na.rm = TRUE, size = 0.4) +
-  geom_line(aes(year, rolMin, colour = '50-yr min'), na.rm = TRUE, size = 0.4) +
-  geom_line(aes(year, rolMed, colour = '50-yr median'), na.rm = TRUE, size = 0.4) +
   geom_smooth(aes(year, rolMax, colour = '50-yr max'), size = 0.6,
@@
     fill = 'magenta4', alpha = 0.1) +
-  geom_col(aes(year, dP, fill = dP)) +
+  geom_col(aes(year, dP, fill = dP), recFinal[dP > 0]) +
   scale_fill_distiller(
-    palette = 'BrBG', limits = abs_range(recFinal$dP),
+    palette = 'BrBG',
+    limits = abs_range(recFinal$dP),
+    # limits = abs_range(recFinal$dP),
+    guide = guide_none(),
+    direction = 1) +
+  ggnewscale::new_scale_fill() +
+  geom_col(aes(year, dP, fill = dP), recFinal[dP < 0]) +
+  scale_fill_distiller(
+    palette = 'BrBG', 
+    # limits = abs_range(recFinal$dP),
+    limits = c(min(recFinal$dP), -min(recFinal$dP)),
     guide = guide_none(),
@@
     size = 2, hjust = 0.8, nudge_y = -15) +
-  # scale_x_continuous(
-  #   breaks = seq(2, 20, 2),
-  #   guide = guide_prism_offset_minor()) +
-  # scale_y_continuous(
-  #   breaks = seq(-300, 0, 50),
-  #   limits = c(-275, -25),
-  #   guide = guide_prism_offset_minor()) +
+  scale_x_continuous(
+    breaks = seq(2, 20, 2),
+    guide = guide_prism_offset_minor()) +
+  scale_y_continuous(
+    breaks = seq(-300, 0, 50),
+    limits = c(-275, -25),
+    guide = guide_prism_offset_minor()) +
   scale_color_manual(

### 2022-08-03 paper-code.Rmd
Add stepwise selection and clean up

@@
 source('R/init.R')
-source('R/input-selection-functions.R')
 source('R/correlation_functions.R')
@@
 
-sv <- c(1, 2)
-sPC <- as.data.table(PC[, sv, drop = FALSE])
+# Stepwise linear regression
+DT <- cbind(PC[450:502, 1:3], Qa = PSepAug$Qa) |> as.data.frame()
+step(lm(Qa ~ ., DT), direction = 'backward')
+```
+
+```{r}
+# Build final model
+sPC <- as.data.table(PC[, 1:2])
 lmFit <- PCR_reconstruction(PSepAug, sPC, firstYear, transform = 'none')
@@
 ```{r}
-round(lmCV$metrics, 2)
+round(lmCV$metrics[, 1:3], 2)
 ```

### 2022-08-03 R/input-selection-functions.R

@@ -1,123 +1,0 @@
-#' Weighted PCA based on correlation, a la Cook et al (2010). Each column of X will be weighted by an exponent r^p, i.e.
-#' \eqn{W_i = X_i*r_i^p}
-#' @param X A matrix of variables
-#' @param rho Vector of correlations between each variable and the output, ignored if `rho == 1`
-#' @param p Weight exponent
-#' @param min.variance Minimum fraction of variance explained by the retained PCs
-#' @param use.eigen IF TRUE, only PCs whose eigenvalue > 1 are kept.
-#' @param return.matrix If `TRUE`, a matrix is returned, otherwise a `data.table` is returned.
-#' @export
-wPCA <- function(X, rho = 1, p = 0, min.variance = 0.95, use.eigen = TRUE, return.matrix = FALSE) {
-
-  # Step 1: weighted PCA
-  # Negative correlation will cause a problem when raised to a real power
-  if (length(rho) > 1) X <- sapply(1:dim(X)[2], function(k) X[, k]*abs(rho[k])^p)
-  # According to Cook et al (2010) SI, when p = 0 we do PCA with correlation matrix form,
-  # and when p != 0 we do PCA with covariance matrix form.
-  pca.model <- summary(stats::prcomp(X, scale. = (p == 0)))
-
-  # Step 2: retain PCs
-  n <- which(pca.model$importance['Cumulative Proportion', ] > min.variance)[1]
-  if (use.eigen) {
-    n2 <- max(which(pca.model$sdev >= 1))
-    if (n2 < n) n <- n2
-  }
-  # If there is only one column, '[' returns a vector instead of a matrix if drop = TRUE
-  out <- pca.model$x[ , 1:n, drop = FALSE]
-  if (return.matrix) out else data.table(out)
-}
-
-#' Input variable selection
-#'
-#' Four methods are in use: Variable Selection Using Random Forests (VSURF) from the `VSURF` package, forward stepwise selection from the `FWDselect` package, and forward and backward selection using branch and bound from the `leaps` package. The latter three use BIC as the selection criterion.
-#' @param X Input variables, can be a matrix or a data.frame. One variable per column. Must have column names.
-#' @param Y Output variable
-#' @param method Either "VSURF", "FWDselect", leaps forward", or "leaps backward".
-#' @param nvmax Maximum number of variables; not applicable for the VSURF method.
-#' @param parallel Only applicable if VSURF or FWDselect is used. Default = `FALSE`.
[…]

### 2022-08-03 paper-code.Rmd
Minor text edits.

@@
 
-First we infill the tree ring data
+First we infill the tree ring data.
 
@@
 
-Calculate bootstrapped correlations
+Calculate bootstrapped correlations.
 
@@
 
+**Build final model**
+
 ```{r}
-# Build final model
 sPC <- as.data.table(PC[, 1:2])
@@
 
-Performance scores
+**Performance scores**
 
@@
 
-Final reconstruction with bias correction
+Final reconstruction with bias correction.
 
@@
 
-Calculate rolling statistics and extract drought events
+Calculate rolling statistics and extract drought events.
 
@@
 
-Mann-Kendall trend test with trend-free prewhitening
+Mann-Kendall trend test with trend-free pre-whitening.
 
@@
 
-Export the reconstruction
+Export the reconstruction.
 

### 2022-10-21 paper-code.Rmd
Revised after peer review.

@@
 library(qmap)         # Bias correction
-library(modifiedmk) # Trend analysis
+library(quantreg)     # Quantile regression
 library(ggplot2)      # Plotting
@@
 library(ggnewscale)   # Plotting
+library(RColorBrewer) # Color palette
 theme_set(theme_prism(base_size = 10, base_fontface = 'plain', base_line_size = 0.2))
@@
 ```{r, eval=FALSE}
-rmarkdown::render('paper-code.Rmd', output_format = 'all', output_options = list(hightlight = 'tango'))
+rmarkdown::render(
+  'paper-code.Rmd', 
+  output_format = 'all', 
+  output_options = list(hightlight = 'tango', 
+                        latex_engine = 'xelatex'))
 ```
@@
 
+### Precipitation and Temperature
+
 ```{r climate data, warning=FALSE, message=FALSE}
@@
 TmWide <- fread('data/chitral-monthly-T.csv')
-Tm <- melt(TmWide, id.var = 'year', variable.name = 'month2', variable.factor = TRUE, value.name = 'Tm'
+Tm <- melt(TmWide, 
+           id.var = 'year', 
+           variable.name = 'month2', 
+           variable.factor = TRUE, 
+           value.name = 'Tm'
          )[, month := as.integer(month2)
@@
 PmWide <- fread('data/chitral-monthly-P.csv')
-Pm <- melt(PmWide, id.var = 'year', variable.name = 'month2', variable.factor = TRUE, value.name = 'Pm'
+Pm <- melt(PmWide, 
+           id.var = 'year', 
+           variable.name = 'month2', 
+           variable.factor = TRUE, 
+           value.name = 'Pm'
          )[, month := as.integer(month2)
@@
 
+### Wind vectors
+
+Note: reanalysis and GPCP data are huge so they won't be included in this repository. You can download them from NOAA in order to run the code below, which produces Figure 1d.
+
+```{r, eval=FALSE}
+# Set map limit
+xrange <- c(0, 80)
+yrange <- c(0, 50)
+
+world <- map('world', fill = TRUE, plot = FALSE) |> 
+  st_as_sf(coords = c('x', 'y')) |> 
+  st_buffer(0.01) |>
+  st_union()
+
+pak <- map('world', 'Pakistan', fill = TRUE, plot = FALSE) |> 
+  st_as_sf(coords = c('xy', 'y'))
+
+## NCEP wind data
+# Wind vector: read u-wind and v-wind, then merge
+# Change the file path to your own
+ncep <- merge(
+  # u-wind
+  ReadNetCDF('data/uwnd.mon.mean.nc', 'uwnd', 
+             subset = list(level = 850, 
+                           lon = xrange, lat = yrange,
+                           time = c('1979-01-01', '2021-12-01'))),
+  ReadNetCDF('data/vwnd.mon.mean.nc', 'vwnd', 
+             subset = list(level = 850, 
+                           lon = xrange, lat = yrange,
+                           time = c('1979-01-01', '2021-12-01'))),
+  by = c('time', 'level', 'lat', 'lon'))
+# Create seasonal mean. 
+ncepSeas <- ncep[month(time) %in% c(2:4, 6:8), 
+                 .(uwnd = mean(uwnd), vwnd = mean(vwnd)), 
+                 by = .(lon, lat, seas = fifelse(month(time) %in% 2:4, 'Feb-Apr', 'Jun-Aug'))]
+
+## GPCP precipitation data
+pre <- ReadNetCDF('data/precip.mon.mean.nc', 'precip',
+                  subset = list(lon = xrange, lat = yrange,
+                           time = c('1979-01-01', '2015-12-01')))
+# Create seasonal mean. 
+preSeas <- pre[month(time) %in% c(2:4, 6:8), 
+               .(precip = mean(precip)), 
+               by = .(lon, lat, seas = fifelse(month(time) %in% 2:4, 'Feb-Apr', 'Jun-Aug'))]
+```
+
+```{r wind plot, fig.width=7, fig.height=3.5, eval=FALSE}
+maxSpeed <- ncepSeas[, sqrt(uwnd^2 + vwnd^2) |> max() |> ceiling()]
+maxRain <- preSeas[, ceiling(max(precip))]
+p3 <- ggplot() +
+  # Precip
+  geom_contour_fill(
+    aes(lon, lat, z = precip, fill = stat(level_d)),
+    data = preSeas,
+    breaks = seq(0, maxRain, 2)) +
+  # Land mass
+  geom_sf(data = world, size = 0.2, fill = NA) +
+  geom_sf(data = pak, size = 0.4, fill = NA, color = 'firebrick') +
+  # Arrow vector
+  geom_arrow(aes(lon, lat, dx = uwnd, dy = vwnd), 
+             data = ncepSeas,
+             # min.mag = 1,
+             # skip = 1,
+             color = blues9[7],
+             size = 0.2,
+             arrow.length = 0.3) +
+  # Site locations
+  geom_point(aes(Longitude, Latitude), crnMeta, color = 'darkviolet') +
+  # Facet by seasons
+  facet_wrap(vars(seas)) +
+  # Set scales
+  as.discretised_scale(scale_fill_distiller)(
+    name = 'Precipitation rate [mm/day]',
+    palette = 'Greens', direction = 1, breaks = seq(0, maxRain, 2),
+    guide = guide_colorsteps(
+      # override.aes = list(alpha = 0.5),
+      title.position = 'top', 
+      show.limits = TRUE)) + 
+  scale_x_continuous(expand = c(0, 0), labels = paste_long) +
+  scale_y_continuous(expand = c(0, 0), labels = paste_lat) +
+  scale_mag(
+    name = 'Wind speed [m/s]', 
+    guide = guide_vector(title.position = 'top'),
+    max = 20) +
+  # guides(alpha = guide_none()) +
+  # Axis labels
+  labs(x = NULL, y = NULL, tag = 'd)') +
+  # Map projection
+  coord_sf(xlim = xrange + c(1, -1.25), ylim = yrange + c(1.25, 0)) +
+  panel_border('black', 0.2) +
+  theme(
+    strip.text = element_text(face = 'bold'),
+    legend.position = 'bottom',
+    legend.key.width = unit(1.25, 'cm'),
+    legend.title = element_text()) 
+p3
+# ggsave('precip-wind-pakistan.png', width = 7, height = 3.5)
+```
+
 ## Tree ring data
@@
 
-**Figure S1 - Chronology and SSS plot**
+**Figure S2 - Chronology and SSS plot**
 
@@
 
-**Figure S2**
+**Figure S5**
 
@@
   melt(id.vars = 'var', variable.name = 'PC')
+
+# Extract PC1 and PC2 for plotting
+pc12plot <- as.data.table(PC[, 1:2, drop = FALSE])[, year := crnWide$year]
 ```
 
-**Figure S3**
+**Figure S6**
 
-```{r, fig.width=7, fig.height=4}
+```{r, fig.width=7, fig.height=6}
 p1 <- ggplot(pcDT[var == 'Standard deviation']) +
@@
 
-p1 + p2
+p3 <- ggplot(pc12plot) +
+  geom_line(aes(year, PC1), color = blues9[9]) +
+  scale_x_continuous(
+    expand = c(0, 0),
+    minor_breaks = seq(1500, 2025, 25),
+    limits = c(1500, 2025),
+    guide = guide_prism_minor()) +
+  labs(x = NULL, tag = 'c)')
+
+p4 <- ggplot(pc12plot) +
+  geom_line(aes(year, PC2), color = blues9[9]) +
+  scale_x_continuous(
+    expand = c(0, 0),
+    minor_breaks = seq(1500, 2025, 25),
+    limits = c(1500, 2025),
+    guide = guide_prism_minor()) +
+  labs(x = NULL, tag = 'd)')
+
+# layout <- '
+# AB
+# CC
+# DD
+# '
+
+layout <- c(
+  patchwork::area(1, 1, 3, 1),
+  patchwork::area(1, 2, 3, 2),
+  patchwork::area(4, 1, 5, 2),
+  patchwork:: area(6, 1, 7, 2)
+)
+p1 + p2 + p3 + p4 +
+  plot_layout(design = layout)
 ```
@@
 step(lm(Qa ~ ., DT), direction = 'backward')
+```
+
+**Figure S1 - Autocorrelation**
+
+```{r}
+acf(PSepAug$Qa, main = NA, xlab = 'Lag [years]')
 ```
@@
 ```
-**Figure S4 - score distribution**
+**Figure S7 - score distribution**
 
@@
 
-Calculate rolling statistics and extract drought events.
+## Quantile regression
 
 ```{r}
-alignType <- 'right'
-recFinal[, rolMax := frollapply(Qbc, 50, max, align = alignType)]
-recFinal[, rolMin := frollapply(Qbc, 50, min, align = alignType)]
-recFinal[, rolMed := frollapply(Qbc, 50, median, align = alignType)]
+taus <- c(0.01, 0.05, 0.50, 0.95, 0.99)
+# names(taus) <- sprintf(taus, fmt = '%.2f')
+names(taus) <- paste0(taus * 100, '%')
+qrDT <- taus |> 
+  lapply(\(tau) {
+    fit <- rq(Qbc ~ year, tau, recFinal)
+    summ <- summary(fit, se = 'boot')
+    data.table(
+      b = fit$coefficients[1], 
+      a = fit$coefficients[2], 
+      signif = summ$coefficients[2, 4] < 0.1)
+  }) |> 
+  rbindlist(idcol = 'tau')
+```
+
+```{r}
 recFinal[, period := fcase(
@@
   list(x = d$x, y = d$y / max(d$y) * 1.5)
-}, by = period]
-
-medians <- recFinal[, {
-  d <- density(Qbc, cut = 0, bw = 45)
-  m <- median(Qbc)
-  y <- approx(d$x, d$y, m)$y
-  list(x = m, y = y / max(d$y) * 1.5)
 }, by = period]
@@
 
-Mann-Kendall trend test with trend-free pre-whitening.
-
-```{r}
-recFinal[!is.na(rolMax), tfpwmk(rolMax)] |> round(6)
-recFinal[!is.na(rolMin), tfpwmk(rolMin)] |> round(6)
-recFinal[!is.na(rolMed), tfpwmk(rolMed)] |> round(6)
-```
-
 **Figure 4**
@@
 ```{r, fig.width=8, fig.height=6, message=FALSE}
-p1 <- ggplot(recFinal) +
-  labs(x = NULL, y = 'P [mm]') +
-  geom_line(aes(year, rolMax, colour = '50-yr max'), na.rm = TRUE, size = 0.4, alpha = 0.5) +
-  geom_line(aes(year, rolMin, colour = '50-yr min'), na.rm = TRUE, size = 0.4, alpha = 0.5) +
-  geom_line(aes(year, rolMed, colour = '50-yr median'), na.rm = TRUE, size = 0.4, alpha = 1) +
-  geom_line(aes(year, Qbc, color = 'Annual')) +
-  geom_smooth(aes(year, rolMax, colour = '50-yr max'), size = 0.6,
-              formula = 'y ~ x', method = 'lm', na.rm = TRUE, fill = NA) +
-  geom_smooth(aes(year, rolMin, colour = '50-yr min'), size = 0.6,
-              formula = 'y ~ x', method = 'lm', na.rm = TRUE, fill = NA) +
-  geom_smooth(aes(year, rolMed, colour = '50-yr median'), size= 0.6,
-              formula = 'y ~ x', method = 'lm', na.rm = TRUE, fill = NA) +
+reds <- brewer.pal(9, 'Reds')
+p1 <- ggplot() +
+  geom_line(aes(year, Qbc), recFinal, color = 'gray85') +
+  geom_abline(aes(slope = a, intercept = b, color = tau, linetype = signif), 
+              qrDT, show.legend = FALSE) +
+  geom_line(aes(x = 1800, y = 400, color = tau, group = 1), qrDT) +
+  geom_point(aes(year, Qbc), recFinal, color = 'gray50', shape = 1, size = 0.3) +
+  scale_color_manual(
+    values = c(reds[6], reds[3], 'darkviolet', blues9[4], blues9[7])) +
+  scale_linetype_manual(values = c(2, 1)) +
   scale_x_continuous(
@@
   scale_y_continuous(
-    minor_breaks = seq(200, 900, 50),
-    breaks = seq(200, 900, 100),
-    limits = c(200, 900),
-    labels = skip_label(2),
+    breaks = seq(200, 800, 100),
+    minor_breaks = seq(200, 850, 50),
+    limits = c(200, 850),
     guide = guide_prism_offset_minor()) +
-  labs(x = NULL, y = 'Precipitation [mm]', colour = NULL, tag = 'a)') +
-  scale_color_manual(values = c(RColorBrewer::brewer.pal(3, 'Set2'), 'gray')) +
+  guides(color = guide_legend(
+    title.position = 'top',
+    title.hjust = 0.5,
+    override.aes = list(size = 0.6))) +
+  labs(x = NULL, y = 'Precipitation [mm]', tag = 'a)') +
   theme(
-    legend.key.width = unit(0.5, 'cm'),
+    plot.tag.position = c(0, 0.94),
     legend.position = 'top')
@@
   geom_line(aes(x, y + period, group = factor(period))) +
-  geom_linerange(aes(x, ymin = period, ymax = period + y), medians, colour = 'firebrick') +
   scale_fill_manual(
@@
 
+# Sensitivity analyses
+
+We now rerun the reconstruction procedure with different SSS cut-off values.
+
+## SSS = 0.75
+
+```{r}
+firstYear2 <- crn[sss > 0.75][, .N, by = year][N >= 4, first(year)]
+lastYear <- 2018
+crnWide2 <- crn[year %in% firstYear2:lastYear,
+                dcast(.SD, year ~ site, value.var = c('res'))]
+X2 <- crnWide2[, -'year'] |> 
+  as.matrix() |> 
+  imputePCA(ncp = 6) |> 
+  {\(x) x$completeObs}() 
+PC2 <- prcomp(X2, scale. = TRUE)$x
+sv <- c(1, 2)
+sPC2 <- as.data.table(PC2[, sv, drop = FALSE])
+lmFit2 <- PCR_reconstruction(PSepAug, sPC2, firstYear2, transform = 'none')
+set.seed(42)
+lmCV2 <- cvPCR(PSepAug, sPC2, firstYear2, transform = 'none',
+              Z = cvFolds, use.robust.mean = FALSE)
+recFinal2 <- lmFit2$rec
+bcFit2 <- fitQmap(
+  PSepAug$Qa,
+  recFinal2[year %in% PSepAug$year, Q],
+  'RQUANT',
+  wet.day = FALSE)
+recFinal2[, Qbc := doQmap(Q, bcFit)]
+```
+
+## SSS = 0.85
+
+```{r}
+firstYear3 <- crn[sss > 0.85][, .N, by = year][N >= 4, first(year)]
+lastYear <- 2018
+crnWide3 <- crn[year %in% firstYear3:lastYear,
+                dcast(.SD, year ~ site, value.var = c('res'))]
+X3 <- crnWide3[, -'year'] |> 
+  as.matrix() |> 
+  imputePCA(ncp = 6) |> 
+  {\(x) x$completeObs}() 
+PC3 <- prcomp(X3, scale. = TRUE)$x
+sv <- c(1, 2)
+sPC3 <- as.data.table(PC3[, sv, drop = FALSE])
+lmFit3 <- PCR_reconstruction(PSepAug, sPC3, firstYear3, transform = 'none')
+lmCV3 <- cvPCR(PSepAug, sPC3, firstYear3, transform = 'none',
+              Z = cvFolds, use.robust.mean = FALSE)
+recFinal3 <- lmFit3$rec
+bcFit3 <- fitQmap(
+  PSepAug$Qa,
+  recFinal3[year %in% PSepAug$year, Q],
+  'RQUANT',
+  wet.day = FALSE)
+recFinal3[, Qbc := doQmap(Q, bcFit)]
+```
+
+## Compare reconstructins from different SSS
+
+**Figure S3**
+
+```{r, fig.width=7, fig.height=6}
+p1 <- ggplot() +
+  geom_line(aes(year, Qbc, color = 'SSS = 0.6', linetype = 'SSS = 0.6'), 
+            recFinal, alpha = 1) +
+  geom_line(aes(year, Qbc, color = 'SSS = 0.75', linetype = 'SSS = 0.75'), 
+            recFinal2, alpha = 1) +
+  # scale_color_manual(values = MetBrewer::met.brewer('Kandinsky', 2))
+  scale_color_manual(name = NULL, 
+                     values = wesanderson::wes_palettes$Cavalcanti1[2:1]) +
+  scale_linetype_manual(name = NULL, values = c(1, 2)) +
+  scale_x_continuous(
+    expand = c(0, 0),
+    minor_breaks = seq(1500, 2025, 25),
+    limits = c(1500, 2025),
+    guide = guide_prism_offset_minor()) +
+  scale_y_continuous(
+    breaks = seq(200, 800, 100),
+    minor_breaks = seq(200, 850, 50),
+    limits = c(200, 850),
+    guide = guide_prism_offset_minor()) +
+  labs(x = NULL, y = 'Precipitation [mm]') +
+  theme(
+    legend.key.height = unit(1, 'cm'),
+    legend.position = 'top')
+p2 <- ggplot() +
+  geom_line(aes(year, Qbc, color = 'SSS = 0.6', linetype = 'SSS = 0.6'), 
+            recFinal, alpha = 1) +
+  geom_line(aes(year, Qbc, color = 'SSS = 0.85', linetype = 'SSS = 0.85'), 
+            recFinal3, alpha = 1) +
+  scale_color_manual(name = NULL, 
+                     values = wesanderson::wes_palettes$Cavalcanti1[c(2,5)]) +
+  scale_linetype_manual(name = NULL, values = c(1, 2)) +
+  scale_x_continuous(
+    expand = c(0, 0),
+    minor_breaks = seq(1500, 2025, 25),
+    limits = c(1500, 2025),
+    guide = guide_prism_offset_minor()) +
+  scale_y_continuous(
+    breaks = seq(200, 800, 100),
+    minor_breaks = seq(200, 850, 50),
+    limits = c(200, 850),
+    guide = guide_prism_offset_minor()) +
+  labs(x = NULL, y = 'Precipitation [mm]') +
+  theme(
+    legend.key.height = unit(1, 'cm'),
+    legend.position = 'top')
+p1 / p2
+```
+
+## Compare skills
+
+**Table S5**
+
+```{r}
+rbindlist(
+  list(
+    s1 = lmCV$metrics,
+    s2 = lmCV2$metrics,
+    s3 = lmCV3$metrics),
+  idcol = 'cutoff')[, 1:4][, cutoff := c(0.6, 0.75, 0.85)][] |> 
+  roundDT(3)
+```
+
+## Compare quantile regression results
+
+```{r}
+taus <- c(0.01, 0.05, 0.50, 0.95, 0.99)
+names(taus) <- paste0(taus * 100, '%')
+qrDT2 <- taus |> 
+  lapply(\(tau) {
+    fit <- rq(Qbc ~ year, tau, recFinal2)
+    summ <- summary(fit, se = 'boot', R = 1000, bsmethod = 'wxy')
+    data.table(
+      b = fit$coefficients[1], 
+      a = fit$coefficients[2], 
+      signif = summ$coefficients[2, 4] < 0.1)
+  }) |> 
+  rbindlist(idcol = 'tau')
+
+qrDT3 <- taus |> 
+  lapply(\(tau) {
+    fit <- rq(Qbc ~ year, tau, recFinal3)
+    summ <- summary(fit, se = 'boot', R = 1000, bsmethod = 'wxy')
+    data.table(
+      b = fit$coefficients[1], 
+      a = fit$coefficients[2], 
+      signif = summ$coefficients[2, 4] < 0.1)
+  }) |> 
+  rbindlist(idcol = 'tau')
+```
+
+**Figure S4**
+
+```{r, fig.width=7, fig.height=6}
+p1 <- ggplot() +
+  geom_line(aes(year, Qbc), recFinal2, color = 'gray85') +
+  geom_abline(aes(slope = a, intercept = b, color = tau, linetype = signif), 
+              qrDT2, show.legend = FALSE) +
+  geom_line(aes(x = 1800, y = 400, color = tau, group = 1), qrDT2) +
+  geom_point(aes(year, Qbc), recFinal2, color = 'gray50', shape = 1, size = 0.3) +
+  scale_color_manual(
+    values = c(reds[6], reds[3], 'darkviolet', blues9[4], blues9[7])) +
+  scale_linetype_manual(values = c(2, 1)) +
+  scale_x_continuous(
+    expand = c(0, 0),
+    minor_breaks = seq(1625, 2025, 25),
+    limits = c(1625, 2025),
+    guide = guide_prism_offset_minor()) +
+  scale_y_continuous(
+    breaks = seq(200, 800, 100),
+    minor_breaks = seq(200, 850, 50),
+    limits = c(200, 850),
+    guide = guide_prism_offset_minor()) +
+  guides(color = guide_legend(
+    title.position = 'top',
+    title.hjust = 0.5,
+    override.aes = list(size = 0.6))) +
+  labs(x = NULL, y = 'Precipitation [mm]') +
+  theme(
+    legend.position = 'top')
+p2 <- ggplot() +
+  geom_line(aes(year, Qbc), recFinal3, color = 'gray85') +
+  geom_abline(aes(slope = a, intercept = b, color = tau, linetype = signif), 
+              qrDT3, show.legend = FALSE) +
+  geom_line(aes(x = 1800, y = 400, color = tau, group = 1), qrDT3) +
+  geom_point(aes(year, Qbc), recFinal3, color = 'gray50', shape = 1, size = 0.3) +
+  scale_color_manual(
+    values = c(reds[6], reds[3], 'darkviolet', blues9[4], blues9[7])) +
+  scale_linetype_manual(values = c(2, 1)) +
+  scale_x_continuous(
+    expand = c(0, 0),
+    minor_breaks = seq(1625, 2025, 25),
+    limits = c(1625, 2025),
+    guide = guide_prism_offset_minor()) +
+  scale_y_continuous(
+    breaks = seq(200, 800, 100),
+    minor_breaks = seq(200, 850, 50),
+    limits = c(200, 850),
+    guide = guide_prism_offset_minor()) +
+  guides(color = guide_legend(
+    title.position = 'top',
+    title.hjust = 0.5,
+    override.aes = list(size = 0.6))) +
+  labs(x = NULL, y = 'Precipitation [mm]') +
+  theme(
+    legend.position = 'top')
+p1 / p2
+```
+
+## Using ARSTAN chronologies
+
+```{r}
+crnWide4 <- crn[year %in% firstYear:lastYear,
+                dcast(.SD, year ~ site, value.var = c('ars'))]
+X4 <- crnWide4[, -'year'] |> 
+  as.matrix() |> 
+  imputePCA(ncp = 6) |> 
+  {\(x) x$completeObs}() 
+PC4 <- prcomp(X4, scale. = TRUE)$x
+sv <- c(1, 2)
+sPC4 <- as.data.table(PC4[, sv, drop = FALSE])
+lmFit4 <- PCR_reconstruction(PSepAug, sPC4, firstYear, transform = 'none')
+lmCV4 <- cvPCR(PSepAug, sPC4, firstYear, transform = 'none',
+              Z = cvFolds, use.robust.mean = FALSE)
+recFinal4 <- lmFit4$rec
+bcFit4 <- fitQmap(
+  PSepAug$Qa,
+  recFinal4[year %in% PSepAug$year, Q],
+  'RQUANT',
+  wet.day = FALSE)
+recFinal4[, Qbc := doQmap(Q, bcFit)]
+```
+
+**Table S3**
+
+```{r}
+rbindlist(list(
+  residual = lmCV$metrics,
+  ARSTAN = lmCV4$metrics),
+  idcol = 'Chronology')[, 1:4] |> 
+  roundDT(3)
+```
+
 Export the reconstruction.
